### 2020-05-03 ArticleAcortado.tex
Acortar

\newcommand{\cc}{\mathfrak{c}}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\CC}{\mathbb{C}}
\newcommand{\Q}{\mathbb{Q}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\Hc}{\mathcal{H}}
\newcommand{\Lan}{\mathcal{L}}
\newcommand{\Ln}{\lim\limits_{n\to \infty}}
\newcommand{\clist}{\mathfrak{c}_{1}, \cdots, \mathfrak{c}_m}
\newcommand{\morph}[1]{\sim_#1}
\newcommand{\vlst}[2]{#1_1,\dots, #1_{#2}}
\newcommand{\gnp}{G(n,\beta_1/n^{a_1-1}, \dots,\beta_l/n^{a_l-1})}
\newcommand{\overbar}[1]{\mkern 1.5mu\overline{\mkern-1.5mu#1\mkern-1.5mu}\mkern 1.5mu}
\newcommand{\ehr}{\textsc{Ehr}}
\newcommand{\PR}[1]{\mathrm{Pr}\big(#1\big)}
\newcommand{\Tr}{\mathrm{Tr}}
\newcommand{\InR}[1]{\left\{ #1_R \right\}_{R\in \sigma}}

@@

Since the work of Erd\"os and R\`enyi on the evolution of random graphs
\cite{erdHos1960evolution} the study of the asymptotic properties of random
structures has played a relevant role in combinatorics and computer science.
A central theme in this topic is, given a succession $(G_n)_n$ of random
structures of some sort and a property $P$, to determine the limit probability
that $G_n$ satisfies $P$ or to determine whether that limit exists. \par
One approach that has proven to be useful is to classify the properties $P$
according to the logical languages they can be defined in. We say that the 
succession $(G_n)_n$ obeys a convergence law with respect to some logical language
$\mathcal{L}$ if for any given property $P$ expressible in $\mathcal{L}$ the 
probability that $G_n$ satisfies $P$ tends to some limit as $n$ grows to infinity.
We say that $(G_n)_n$ obeys a zero-one law with respect to $\mathcal{L}$ if
that limit is always either zero or one. 
The seminal theorem on this topic, due to Fagin \cite{fagin1976probabilities}
and Glebskii et al. 
\cite{glebskii1969range} independently, states that if $G_n$ denotes a labeled
graph with $n$ vertices picked uniformly at random among all $2^{\binom{n}{2}}$ 
possible then $(G_n)_n$ satisfies a zero-one law with respect to the first order
(FO) language of graphs. \par
Originally this result was proven in the broader context of relational
structures but it was in the theory of random graphs were the study of 
other zero-one and convergence laws became more prominent. In particular,
the asymptotic behavior of FO logic in the binomial model of random graphs 
$G(n,p)$ has been extensively studied. In this model, introduced by Gilbert 
\cite{gilbert1959random}, a random graph is obtained from $n$ labeled vertices
by adding each possible edge with probability $p$ independently. When $p=1/2$
this distribution of random graphs coincides with the uniform one, mentioned 
above. In general, for the case where $p$ is a constant probability a slight 
generalization of the proofs in \cite{fagin1976probabilities} and 
\cite{glebskii1969range} works and $G(n,p)$ satisfies a zero-one law 
for FO logic. If we consider $p(n)$ a decreasing function of the form
$n^{-\alpha}$ we can ask the question of what are the values of $\alpha$
for which $G(n,p(n))$ obeys a zero-one or a convergence law for FO logic. 
In \cite{shelah1988zero} Shelah and Spencer gave a complete answer 
for the range $\alpha\in (0,1)$. Among other results, they proved that
if $\alpha$ is an irrational number in this interval then
$G(n,p(n))$ obeys a zero-one law for FO logic, while if $\alpha$
is a rational number in the same range then $G(n,p(n))$ does not
even satisfy a convergence law for FO logic. The case $\alpha=1$
was later solved by Lynch in \cite{lynch1992probabilities}. A weaker
form of the main theorem in that article states the following:
\begin{theorem} 
	For any FO sentence $\phi$, the function
	$F_\phi: (0,\infty)\rightarrow [0,1]$ given by 
	\[ F_\phi(\beta) = \Ln \mathrm{Pr}\big( \mathrm{G}(n,\beta/n)
	\text{ satisfies } \phi   \big) \]
	is well defined and analytic. In particular, for any
	$\beta \geq 0$ the model $\mathrm{G}(n,\beta/n)$
	obeys a convergence law for FO logic. 
\end{theorem}

@@

Given a positive natural number $n$, we will write
$[n]$ to denote the set ${1,2,\dots,n}$.\par
Given a set $S$ and a natural number $k\in \N$
we will use $\binom{S}{k}$ to denote the set of 
subsets of $S$ whose size is $k$. \par
Given numbers, $n,m\in \N$ with $m\leq n$ we define
$(n)_m:=n\cdot (n-1)\cdots (n-m+1)$. Given a set
$S$ and a number $n\in \N$ with $n\leq |S|$ we define
$(S)_n$ as the subset of $S^n$ consisting of the $n$-tuples
whose coordinates are all different. 
We also define $S^*:=\bigcup_{n=0}^\infty S^n$ and
$(S)_*=\bigcup_{n\leq |S|} (S)_n$. Given a tuple 
$\overline{x}\in S^*$ and an element $x\in S$ the expression
$x\in \overline{x}$ will mean that $x$ appears as some coordinate
in $\overline{x}$. We will at times make an abuse of notation and
treat the tuple $\overline{x}$
as the set of elements $x\in \overline{x}$. \par
We will use the convention that over-lined variables
, like $\overline{x}$, denote ordered tuples of arbitrary length. 
For example, given a set $S$, if we write $\overline{x}\in S^*$
then $\overline{x}=(x_1,\dots, x_a)$ for some $a\in \N$ and 
some $x_1,\dots, x_a\in S$. Given an ordered tuple $\overline{x}$
we define the number $len(\overline{x})$ as its length. 
Given a map $f:X\rightarrow Y$ between two sets $X, Y$ and 
an ordered tuple $\overline{x}:=(x_1,\dots,x_a)\in X^*$ 
we define $f(\overline{x})\in Y^*$ as the tuple 
$(f(x_1),\dots,f(x_a))$.
\par

Let $S$ be a set, $a$ a positive natural number, 
and $\Phi$ a group of permutations over 
$[a]$. Then $\Phi$ acts naturally over
$S^a$ in the following way: Given $g\in \Phi$ and
$\overline{x}=(x_1,\dots,x_a)\in S^a$ we define 
$g\cdot(x_1,\dots,x_a)$ and $g\cdot \overline{x}$ 
as the tuple $(x_{g(1)},\dots,x_{g(a)})$. 
We will denote by $S^a/\Phi$ to the quotient
of the set $S^a$ by this action. Given an element
$\overline{x}:=(x_1,\dots, x_a)\in S^a$ we will denote its equivalence
class in $S^a/\Phi$ by $[x_1,\dots,x_a]$ or $[\overline{x}]$.
Thus, for any
$g\in \Phi$, by definition $[x_1,\dots,x_a]=[x_{g(1)}
,\dots,x_{g(a)}]$. The notations $\overline{x}$ and
$(x_1,\dots, x_a)$ 
will be reserved 
to ordered tuples while 
$[\overline{x}]$ and
$[x_1,\dots,x_a]$ will denote ordered tuples modulo the
action of some arbitrary group of permutations. Which group is
this will depend on the ambient set where $[x_1,\dots,x_a]$ belongs
and it should either be clear from context or not be relevant.\par
\par
Given two real functions over the natural numbers 
$f,g:\N \rightarrow \R$ we will write $f=O(g)$ to 
mean that there exists some constant $C\in \R$
such that $f(n)\leq Cg(n)$ for $n$ sufficiently large, 
as usual.
If $g(n)\neq 0$ for $n$ large enough then
we will write $f\sim g$ when $\Ln \frac{f(n)}{g(n)}=1$.\par

\todo[inline]{Añadir referencias y trabajo relacionado con random SAT}

\subsection{Logical preliminaries}
We will assume a certain degree of familiarity with the 
concepts. For a more complete exposition of the topics 
presented here one can consult \cite{ebbinghaus2013mathematical}.

A relational vocabulary $\sigma$ is a collection of
relation symbols $\{R_1,\dots, R_m,\dots\}$
where each relation symbol $R_i$ has associated a natural
number $ar(R_i)$ called its arity.
A $\sigma$-structure $\mathfrak{A}$ is composed of
a set $A$, called the universe of $\mathfrak{A}$, equipped with
relations $R_1^{\mathfrak{A}}\subseteq A^{a_1},
\dots, R_m^{\mathfrak{A}}\subseteq A^{a_m}$. When $\sigma$
is understood we may refer to $\sigma$-structures as relational
structures or simply as structures.
A structure is called finite if its universe is a finite set. \par

In the first order language $FO[\sigma]$ with signature $\sigma$ 
formulas are formed by variables $x_1,\dots,$ $x_i,\dots$,
the relation symbols in $\sigma$, the equal symbol $=$,
the usual Boolean connectives $\neg, \wedge, \vee, \dots$, 
the existential and universal quantifiers $\exists, \forall$, and
the parentheses $),($. Then formulas in $FO[\sigma]$ are defined as 
follows.
\vspace*{-0.2em}
\begin{itemize}[topsep=1pt, itemsep=1pt]
	\item The expression $R(y_1,\dots,y_a)$, where the $y_i$'s are variables and $R$
	is a relation symbol in $\sigma$ such that $ar(R)=a$,
	belongs to $FO[\sigma]$.
	\item The expression $y_1=y_2$, where $y_1, y_2$ are variables, 
	belongs to $FO[\sigma]$.
	\item Given formulas $\phi, \psi \in FO[\sigma]$, any Boolean combination of them
	$\neg(\phi), (\phi \wedge \psi), (\phi \vee \psi), \dots$ belongs to $FO[\sigma]$ as well.
	\item Given a formula $\phi \in FO[\sigma]$ and $x$ a variable that 
	does not appear bounded
	by a quantifier in $\phi$, the expressions $\forall x (\phi)$ and 
	$\exists x (\phi)$ belong both to $FO[\sigma]$. 
\end{itemize}  
We will write $\forall y_1, y_2,\dots, y_m$ or simply $\forall \overline{y}$
instead of  
$\forall y_1, \forall y_2,\dots, \forall y_m$ and likewise for
the quantifier $\exists$. Also, if $\overline{y}:=(y_1,\dots,y_a)$ is a
tuple of variables we may write simply $R(\overline{y})$ instead
of $R(y_1\dots,y_a)$.\par
For the remaining of this article we will reserve the names $x,y,z$
for the variables in our first order formulas.\par
We define the set of free variables of a formula as usual. 
Given a formula $\phi\in FO[\sigma]$ we will use the notation $\phi(\overline{y})$ 
to denote that $\overline{y}$ is a tuple of 
(different) variables that contains all free variables in $\phi$ and
none of its bounded variables, although it may contain variables
which not appear in $\phi$. \par
Formulas with no free variables are called sentences and 
formulas with no quantifiers are called open formulas. \par
The quantifier rank of a formula $\phi$, denoted by $qr(\phi)$, 
is defined as the maximum number of nested quantifiers in $\phi$. \par
Sentences in $FO[\sigma]$ are interpreted over $\sigma$-structures
in the natural way. Given an structure $\mathcal{A}$, and a sentence
$\phi\in FO[\sigma]$ we write $\mathcal{A} \models \phi$ to denote 
that $\mathcal{A}$ satisfies $\phi$. If $\psi(\overline{y})$ is a
formula, $\overline{a}$ are elements in the universe of 
$\mathcal{A}$, and $\overline{y}$ and $\overline{a}$ are lists of the same size,
then we write $\mathcal{A} \models 
\psi(\overline{a})$
to mean that $\mathcal{A}$ satisfies $\psi$ when the free variables in
$\overline{y}$ are interpreted as the elements in $\overline{a}$ (i.e., 
the $i$-th element of $\overline{y}$ is interpreted as the $i$-th
element of $\overline{a}$).

\subsection{Structures as multi-hypergraphs}

@@

We will only consider relational structures where the relations
are of arity at least two. This restriction is not necessary, but 
it makes notation easier. 
\todo[inline]{Quiz\'as deber\'ia a\~nadir un anexo dando alguna indicaci\'on sobre c\'omo tratar las relaciones unarias?}

@@

We can think any structure $G$ in $\mathcal{C}$
as a "multi-hypergraph" whose vertices are the elements
of the universe of $G$. Each relation $R\in \sigma$, can be 
represented over $G$ as an "edge set" formed by tuples vertices
of size $ar(R)$ modulo the action of $\Phi_R$. Furthermore, 
repetitions of vertices in the positions given by
$P_R$ are not allowed in these tuples.\par
The following definitions make this ideas formal. They 
depend on our choices of $\sigma$, $\{\phi_R\}_{R\in \sigma}$
and $\{P_R\}_{R\in \sigma}$ but as those are fixed we can
allow ourselves to omit those dependencies for the sake of readability.

@@

Hypergraphs, as we have defined them, 
can be naturally interpreted as
structures from $\mathcal{C}$ 
in the following way: 
given $G=(V,\{E_R\}_{R\in \sigma})$,
we consider $V$ to be the universe of
 $G$, and for any $R\in \sigma$ we define
$R^G\subseteq V^{ar(R)}$ as the set of tuples
$\overline{x}\in V^{ar(R)}$ such that
$[\overline{x}]\in E_R$. Under 
this interpretation hypergraphs
satisfy by definition the 
symmetry and anti-reflexivity
axioms given at the beginning of this section. It is also easy to
see that this interpretation induces
a one-to-one identification between 
structures in $\mathcal{C}$ and hypergraphs. \par

@@

We will use standard nomenclature and notation
from graph theory with some additions. Given an 
hypergraph $G$ we will call its \textbf{vertex set} to $V(G)$ 
and \textbf{vertices} 
to the elements $v\in V(G)$. Likewise, each of 
the $E_R(G)$'s will be called an \textbf{edge set} and its elements, 
\textbf{edges}. Given an edge set $E_R(G)$, 
the index $R$ will be called its \textbf{relation}.

@@

Given a set of vertices $U\subseteq V(G)$, 
we will denote by $G[U]$ the \textbf{hypergraph induced
by $G$ on $U$}. That is, $G[U]$ is an hypergraph
$H=(V(H),\{E(H)_R\}_{R\in \sigma})$ such that 
$V(H)=U$ and for any $R\in \sigma$ 
each edge $e\in E_R(G)$ belongs 
to $E_R(H)$ as well if and only if the vertices
involved in $e$ are in $U$ (i.e. $e\in E_R[U]$).
\par
%
%An homomorphism between two hypergraphs $G$ and $H$ 
%a map $f:V(G)\rightarrow V(H)$ that sends edges 
%from $G$ to edges in $H$ of the same color. That is,
%if vertices $v_1,\dots, v_{a_i}$ form an edge 
%$[v_1,\dots,v_{a_i}]\in E_i(G)$, then
%$[f(v_1),\dots,f(v_{a_i})]\in E_i(H)$.
%If $f$ is injective then it is called a monomorphism. 
%If $f$ is bijective and its inverse is also an homomorphism 
%between $H$ and $G$ then $f$ is called an isomorphism.\par
%
%The group of automorphisms $Aut(G)$ of an hypergraph $G$ is the group
%of isomorphisms between $G$ and itself.

%
%Given two hypergraphs $G$ and $H$, a copy of $H$ in $G$ is
%a sub-hypergraph $H_2\subseteq G$ isomorphic to $H$. The copy
%is called induced if $H_2$ is an induced sub-hypergraph.
%We will call a labeled copy
%of $H$ in $G$ to a monomorphism $f:H\rightarrow G$. 
%It is satisfied that the number
%of labeled copies of $H$ in $G$ is $|Aut(H)|$ times the number
%of copies of $H$ in $G$.

We define the \textbf{excess} $ex(G)$ of an hypergraph $G$ as the number
\[
ex(G):= \big(\sum_{R\in \sigma} (ar(R)-1)|E_R(G)|\big) - |V(G)|.  
\] 
That is, the excess of $G$ is its "weighted number of edges"
minus its number of vertices. \par

Given an hypergraph $G$ we define the following metric, $d$, over $V(G)$:
\[ d^G(u,v)= \min_{\substack{H \subset G\\ 
		H \text{ connected }\\
		u,v\in V(H)}} |E(H)| .\]
That is, the distance between $v$ and $u$ is the minimum number of
edges necessary to connect $v$ and $u$. 
If such number does not exist we define $d^G(u,v)=\infty$. \par
As usual, define $d^{G}(X,Y)$ for sets $X,Y\subset V(G)$ as the minimum
distance $d(u,v)$ where $u\in X$ and $v\in Y$. 
When $X=\{x\}$ we will omit the brackets and write
$d^G(x,Y)$ instead of $d^G(\{x\},Y)$, for example.
When $G$ 
is understood or not relevant we will usually simply denote the 
distance by $d$ instead of $d^G$. \par

@@

Given two hypergraphs $H_1, H_2$ an \textbf{isomorphism} 
$f$ between them is a bijection $f: V(H_1)\rightarrow V(H_2)$
such that for any $R\in \sigma$, a tuple $\overline{v}\in V(H_1)^*$
forms an edge $[\overline{v}]\in E_\sigma(H_1)$ if and only
if $[f(\overline{v})]\in E_\sigma(H_2)$. We say that $H_1$ and
$H_2$ are isomorphic, written as $H_1\simeq H_2$, if there exists
some isomorphism between them.\par

Let $H$ be an hypergraph. An \textbf{automorphism} of $H$
is an isomorphism from $H$ to itself. We will denote by
$Aut(H)$ the group of such automorphisms. \par

Let $H^\simeq$ be an isomorphism class of hypergraphs. Let
$H\in H^\simeq$ be a representative of the class. 
Then we define $ex(H^\simeq)$ as $ex(H)$. Clearly this definition
does not depend on the choice of $H$. The expressions
$|Aut(H^\simeq)|$,
$|V(H^\simeq)|$, $|E(H^\simeq)|$ or
$|E_R(H^\simeq)|$ for $R\in \sigma$ are defined 
in an analogous way.\par

Let $H$ be an hypergraph and let $V$ be a set. We define the
set $Copies(H,V)$ as the set of hypergraphs $H^\prime$ such that
$V(H^\prime)\subset V$ and $H\simeq H^\prime$. Analogously, if
$H^\simeq$ is an isomorphism class of hypergraphs we define
$Copies(H^\simeq,V)$ as the set of hypergraphs $H^\prime$ 
that satisfy
$V(H^\prime)\subset V$ as well as $H^\prime \in H^\simeq$.   \par

@@

\begin{definition}
	Let $\Sigma$ be a set. A \textbf{$\Sigma$-hypergraph}
	is a pair $(H, \chi)$ where $H$ is an hypergraph
	and $\chi: V(H)\rightarrow \Sigma$ is a map 
	called \textbf{$\Sigma$-coloring} of $H$.
\end{definition}

>>>>

\newcommand{\cc}{\mathfrak{c}}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\CC}{\mathbb{C}}
\newcommand{\Q}{\mathbb{Q}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\Hc}{\mathcal{H}}
\newcommand{\Lan}{\mathcal{L}}
\newcommand{\Ln}{\lim\limits_{n\to \infty}}
\newcommand{\clist}{\mathfrak{c}_{1}, \cdots, \mathfrak{c}_m}
\newcommand{\morph}[1]{\sim_#1}
\newcommand{\vlst}[2]{#1_1,\dots, #1_{#2}}
\newcommand{\gnp}{G(n,\beta_1/n^{a_1-1}, \dots,\beta_l/n^{a_l-1})}
\newcommand{\overbar}[1]{\mkern 1.5mu\overline{\mkern-1.5mu#1\mkern-1.5mu}\mkern 1.5mu}
\newcommand{\ehr}{\textsc{Ehr}}
\newcommand{\PR}[1]{\mathrm{Pr}\big(#1\big)}
\newcommand{\Tr}{\mathrm{Tr}}
\newcommand{\InR}[1]{\left\{ #1_R \right\}_{R\in \sigma}}
\newcommand{\sep}{\noindent\rule{2cm}{0.4pt}}
\newcommand{\aut}{\mathrm{aut}}
\DeclarePairedDelimiter\floor{\lfloor}{\rfloor}

@@

\todo[inline]{INCOMPLETA, PENDIENTE DE SER REVISADA}
Since the work of Erd\"os and R\`enyi on the evolution of random graphs
\cite{erdHos1960evolution} the study of the asymptotic properties of random
structures has played a relevant role in combinatorics and computer science.
A central theme in this topic is, given a succession $(G_n)_n$ of random
structures of some sort and a property $P$, to determine the limit probability
that $G_n$ satisfies $P$ or to determine whether that limit exists. \par
One approach that has proven to be useful is to classify the properties $P$
according to the logical languages they can be defined in. We say that the 
succession $(G_n)_n$ obeys a convergence law with respect to some logical language
$\mathcal{L}$ if for any given property $P$ expressible in $\mathcal{L}$ the 
probability that $G_n$ satisfies $P$ tends to some limit as $n$ grows to infinity.
We say that $(G_n)_n$ obeys a zero-one law with respect to $\mathcal{L}$ if
that limit is always either zero or one. 
The seminal theorem on this topic, due to Fagin \cite{fagin1976probabilities}
and Glebskii et al. 
\cite{glebskii1969range} independently, states that if $G_n$ denotes a labeled
graph with $n$ vertices picked uniformly at random among all $2^{\binom{n}{2}}$ 
possible then $(G_n)_n$ satisfies a zero-one law with respect to the first order
(FO) language of graphs. \par
Originally this result was proven in the broader context of relational
structures but it was in the theory of random graphs were the study of 
other zero-one and convergence laws became more prominent. In particular,
the asymptotic behavior of FO logic in the binomial model of random graphs 
$G(n,p)$ has been extensively studied. In this model, introduced by Gilbert 
\cite{gilbert1959random}, a random graph is obtained from $n$ labeled vertices
by adding each possible edge with probability $p$ independently. When $p=1/2$
this distribution of random graphs coincides with the uniform one, mentioned 
above. In general, for the case where $p$ is a constant probability a slight 
generalization of the proofs in \cite{fagin1976probabilities} and 
\cite{glebskii1969range} works and $G(n,p)$ satisfies a zero-one law 
for FO logic. If we consider $p(n)$ a decreasing function of the form
$n^{-\alpha}$ we can ask the question of what are the values of $\alpha$
for which $G(n,p(n))$ obeys a zero-one or a convergence law for FO logic. 
In \cite{shelah1988zero} Shelah and Spencer gave a complete answer 
for the range $\alpha\in (0,1)$. Among other results, they proved that
if $\alpha$ is an irrational number in this interval then
$G(n,p(n))$ obeys a zero-one law for FO logic, while if $\alpha$
is a rational number in the same range then $G(n,p(n))$ does not
even satisfy a convergence law for FO logic. The case $\alpha=1$
was later solved by Lynch in \cite{lynch1992probabilities}. A weaker
form of the main theorem in that article states the following:
\begin{theorem} 
	For any FO sentence $\phi$, the function
	$F_\phi: (0,\infty)\rightarrow [0,1]$ given by 
	\[ F_\phi(\beta) = \Ln \mathrm{Pr}\big( \mathrm{G}(n,\beta/n)
	\text{ satisfies } \phi   \big) \]
	is well defined and analytic. In particular, for any
	$\beta \geq 0$ the model $\mathrm{G}(n,\beta/n)$
	obeys a convergence law for FO logic. 
\end{theorem}

@@

Given a positive natural number $n$, we will write
$[n]$ to denote the set ${1,2,\dots,n}$.\par
Given a set $S$ and a natural number $k\in \N$
we will use $\binom{S}{k}$ to denote the set of 
subsets of $S$ whose size is $k$. \par
Given numbers, $n,m\in \N$ with $m\leq n$ we denote by
$(n)_m$ the $m$-th falling factorial of $n$. \par
Given a set
$S$ and a number $n\in \N$ with $n\leq |S|$ we define
$(S)_n$ as the subset of $S^n$ consisting of the $n$-tuples
whose coordinates are all different. 
We also define $S^*:=\bigcup_{n=0}^\infty S^n$ and
$(S)_*:=\bigcup_{n\leq |S|} (S)_n$. 
Given two tuples $\overline{x},\overline{y}\in S^*$
we write $\overline{x}^\smallfrown \overline{y}$ to denote their 
concatenation. Given a tuple 
$\overline{x}\in S^*$ and an element $x\in S$ the expression
$x\in \overline{x}$ will mean that $x$ appears as some coordinate
in $\overline{x}$. We will at times make an abuse of notation and
treat the tuple $\overline{x}$
as the set of elements $x\in \overline{x}$. 
We will use the convention that over-lined variables
, like $\overline{x}$, denote ordered tuples of arbitrary length.
Given an ordered tuple $\overline{x}$
we define the number $len(\overline{x})$ as its length. 
Given a map $f:X\rightarrow Y$ between two sets $X, Y$ and 
an ordered tuple $\overline{x}:=(x_1,\dots,x_a)\in X^*$ 
we define $f(\overline{x})\in Y^*$ as the tuple 
$(f(x_1),\dots,f(x_a))$.
\par

Let $S$ be a set, $a$ a positive natural number, 
and $\Phi$ a group of permutations over 
$[a]$. Then $\Phi$ acts naturally over
$S^a$ in the following way: Given $g\in \Phi$ and
$\overline{x}:=(x_1,\dots,x_a)\in S^a$ we define 
$g\cdot \overline{x}$ 
as the tuple $(x_{g(1)},\dots,x_{g(a)})$. 
We will denote by $S^a/\Phi$ to the quotient
of the set $S^a$ by this action. Given an element
$\overline{x}:=(x_1,\dots, x_a)\in S^a$ we will denote its equivalence
class in $S^a/\Phi$ by $[x_1,\dots,x_a]$ or $[\overline{x}]$.
Thus, for any
$g\in \Phi$, by definition $[x_1,\dots,x_a]=[x_{g(1)}
,\dots,x_{g(a)}]$. The notations $\overline{x}$ and
$(x_1,\dots, x_a)$ 
will be reserved 
to ordered tuples while 
$[\overline{x}]$ and
$[x_1,\dots,x_a]$ will denote ordered tuples modulo the
action of some arbitrary group of permutations. Which group is
this will depend on the ambient set where $[x_1,\dots,x_a]$ belongs
and it should either be clear from context or not be relevant.\par

Given two real functions over the natural numbers 
$f,g:\N \rightarrow \R$ we will write $f=O(g)$ to 
mean that there exists some constant $C\in \R$
such that $f(n)\leq Cg(n)$ for $n$ sufficiently large, 
as usual. We will write $f=\Theta(g)$ if both $f=O(g)$ and
$g=O(f)$. If $g(n)\neq 0$ for $n$ large enough then
we will write $f\sim g$ when $\Ln \frac{f(n)}{g(n)}=1$.

%\todo[inline]{Añadir referencias y trabajo relacionado con random SAT}

\subsection{Logical preliminaries}
We assume familiarity with first order logic (FO). We follow 
the convention that first order logic contains the equality symbol. 
Given a vocabulary $\sigma$ we will denote by $FO[\sigma]$ the set of 
first order formulas of vocabulary $\sigma$. We define the set of 
\textbf{free variables} of a formula as usual. 
Given a relation symbol $R\in \sigma$ we denote by $ar(R)$ the arity of $R$. 
Given a formula $\phi\in FO[\sigma]$ we will use the notation $\phi(\overline{y})$ 
to denote that $\overline{y}$ is a tuple of 
(different) variables that contains all free variables in $\phi$ and
none of its bounded variables, although it may contain variables
which not appear in $\phi$.
Formulas with no free variables are called \textbf{sentences} and 
formulas with no quantifiers are called \textbf{open formulas}.

\subsection{Structures as multi-hypergraphs} \label{sect:structures}

@@

\end{itemize}
%
%We will only consider relational structures where the relations
%are of arity at least two. We claim that it is also possible to prove
%our results for structures that include unary relations but we will refrain from
%doing so because it would add unnecessary complexity to our proofs. \par

@@

We can think any structure $G$ in $\mathcal{C}$
as a ``multi-hypergraph" whose vertices are the elements
of the universe of $G$. Each relation $R\in \sigma$, can be 
represented over $G$ as an ``edge set" formed by tuples vertices
of size $ar(R)$ modulo the permutations in $\Phi_R$. Furthermore, 
repetitions of vertices in the positions given by
$P_R$ are not allowed in these tuples.\par
The following definitions make this ideas formal. Notice that they 
depend on our fixed choice of $\sigma$, $\{\phi_R\}_{R\in \sigma}$
and $\{P_R\}_{R\in \sigma}$.

@@

Hypergraphs, as we have defined them, 
can be naturally interpreted as
structures from $\mathcal{C}$ 
in the following way: 
given an hypergraph $G:=(V,\{E_R\}_{R\in \sigma})$,
we can construct a $\sigma$-structure $\mathcal{G}$ whose universe is
$V$, and for any $R\in \sigma$ we define
$R^\mathcal{G}\subseteq V^{ar(R)}$ as the set of tuples
$\overline{x}\in V^{ar(R)}$ such that
$[\overline{x}]\in E_R$. Clearly $\mathcal{G}$
satisfies by definition the 
symmetry and anti-reflexivity
axioms given at the beginning of this section. It is easy to
see that this process induces
a one-to-one identification between 
structures in $\mathcal{C}$ and hypergraphs. \par

@@

We will use standard nomenclature and notation
from graph theory with some additions. Given an 
hypergraph $G$ we will call its \textbf{vertex set} to $V(G)$ 
and \textbf{vertices} 
to the elements $v\in V(G)$. Likewise, each of 
the $E_R(G)$'s will be called an \textbf{edge set} and its elements, 
\textbf{edges}. Given an edge, $e\in V(G)$ we will denote by $V(e)$
the set of all vertices that participate in $e$.

@@

Given a set of vertices $U\subseteq V(G)$, 
we will denote by $G[U]$ the \textbf{hypergraph induced
by $G$ on $U$}. That is, $G[U]$ is an hypergraph
$H=(V(H),\{E(H)_R\}_{R\in \sigma})$ such that 
$V(H)=U$ and for any $R\in \sigma$ 
an edge $e\in E_R(G)$ belongs 
to $E_R(H)$ if and only if $V(e)\subset U$.
\par
%
%An homomorphism between two hypergraphs $G$ and $H$ 
%a map $f:V(G)\rightarrow V(H)$ that sends edges 
%from $G$ to edges in $H$ of the same color. That is,
%if vertices $v_1,\dots, v_{a_i}$ form an edge 
%$[v_1,\dots,v_{a_i}]\in E_i(G)$, then
%$[f(v_1),\dots,f(v_{a_i})]\in E_i(H)$.
%If $f$ is injective then it is called a monomorphism. 
%If $f$ is bijective and its inverse is also an homomorphism 
%between $H$ and $G$ then $f$ is called an isomorphism.\par
%
%The group of automorphisms $Aut(G)$ of an hypergraph $G$ is the group
%of isomorphisms between $G$ and itself.

%
%Given two hypergraphs $G$ and $H$, a copy of $H$ in $G$ is
%a sub-hypergraph $H_2\subseteq G$ isomorphic to $H$. The copy
%is called induced if $H_2$ is an induced sub-hypergraph.
%We will call a labeled copy
%of $H$ in $G$ to a monomorphism $f:H\rightarrow G$. 
%It is satisfied that the number
%of labeled copies of $H$ in $G$ is $\aut(H)$ times the number
%of copies of $H$ in $G$.

We define the \textbf{excess} $ex(G)$ of an hypergraph $G$ as the number
\[
ex(G):= \big(\sum_{R\in \sigma} (ar(R)-1)|E_R(G)|\big) - |V(G)|.  
\] 
That is, the excess of $G$ is its "weighted number of edges"
minus its number of vertices. \par
An hypergraph $G$ is \textbf{connected} if for any two vertices $v,u\in V(G)$
there is a sequence of edges $e_1,\dots, e_m\in E(G)$ such that
$v\in V(e_1), u\in V(e_m)$ and for each $i\in [m-1]$, 
$V(e_i)\cap V(e_{i+1})\neq \emptyset$. It holds that
$ex(G)\geq -1$ for any connected hypergraph.
\par Given a connected hypergraph $G$ 
and vertices $v,u\in V(G)$, we say that $G$ is
a \textbf{path} between $v$ and $u$
if $G$ 
does not contain any connected proper sub-hypergraph containing both $v,u$.
\par
A connected hypergraph $G$ is a \textbf{tree} if $ex(G)=-1$. 
\par A connected hypergraph $G$ is called \textbf{dense} if $ex(G)>0$.\par
A connected hypergraph $G$ with $ex(G)\geq 0$ is called \textbf{saturated} 
if for any non-empty proper sub-hypergraph
$H\subset G$ it holds $ex(H)<ex(G)$.
% It is satisfied that any connected hypergraph $G$
%with $ex(G)\geq 0$ contains a unique saturated sub-hypergraph $H$ with $ex(H)=ex(G)$.
\par
A connected hypergraph $G$ with $ex(G)=0$ is called \textbf{unicycle}. 
A saturated unicycle is called a \textbf{cycle}. \par
Given an hypergraph $G$ we define the following metric, $d$, over $V(G)$:
\[ d^G(u,v)= \min_{\substack{H \subset G\\ 
		H \text{ connected }\\
		u,v\in V(H)}} |E(H)| .\]
That is, the distance between $v$ and $u$ is the minimum number of
edges necessary to connect $v$ and $u$. 
If such number does not exist we define $d^G(u,v)=\infty$. \par
As usual, define $d^{G}(X,Y)$ for sets $X,Y\subset V(G)$ as the minimum
distance $d(u,v)$ where $u\in X$ and $v\in Y$. 
When $X=\{x\}$ we will omit the brackets and write
$d^G(x,Y)$ instead of $d^G(\{x\},Y)$, for example.
When $G$ 
is understood or not relevant we will usually simply denote the 
distance by $d$ instead of $d^G$. \par

@@

Given two hypergraphs $H_1, H_2$ an \textbf{isomorphism} 
$f$ between them is a bijection $f: V(H_1)\rightarrow V(H_2)$
such that for any $R\in \sigma$, a tuple $\overline{v}\in V(H_1)^*$
forms an edge $[\overline{v}]\in E_R(H_1)$ if and only
if $[f(\overline{v})]\in E_R(H_2)$. We say that $H_1$ and
$H_2$ are isomorphic, written as $H_1\simeq H_2$, if there exists
some isomorphism between them.\par

Let $H$ be an hypergraph. An \textbf{automorphism} of $H$
is an isomorphism from $H$ to itself. We will denote by
$\aut(H)$ the number of such automorphisms. \par

Let, $\mathbb{H}$ be an isomorphism class of hypergraphs. Let
$H\in \mathbb{H}$ be a representative of the class. 
Then we define $ex(\mathbb{H})$ as $ex(H)$. Clearly this definition
does not depend on the choice of $H$. The expressions
$\aut(\mathbb{H})$,
$|V(\mathbb{H})|$, $|E(\mathbb{H})|$ or
$|E_R(\mathbb{H})|$ for $R\in \sigma$ are defined 
in an analogous way.\par

Let $H$ be an hypergraph and let $V$ be a set. We define the
set $Copies(H,V)$ as the set of hypergraphs $H^\prime$ such that
$V(H^\prime)\subset V$ and $H\simeq H^\prime$. Analogously, if
$\mathbb{H}$ is an isomorphism class of hypergraphs we define
$Copies(\mathbb{H},V)$ as the set of hypergraphs $H^\prime$ 
that satisfy
$V(H^\prime)\subset V$ as well as $H^\prime \in \mathbb{H}$.   \par

@@

\begin{definition} \label{def:color}
	Let $\Sigma$ be a set. A \textbf{$\Sigma$-hypergraph}
	is a pair $(H, \chi)$ where $H$ is an hypergraph
	and $\chi: V(H)\rightarrow \Sigma$ is a map 
	called \textbf{$\Sigma$-coloring} of $H$.
\end{definition}
